---
title: "Machine Learning Course Project"
output: html_document
---

# Machine Learning Course Project

## Download Dataset
```{r, eval=FALSE}
trainingURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(trainingURL,"pml-training.csv",method="curl")
download.file(testingURL, "pml-testing.csv", method="curl")
```

## Load Libraries
```{r, results="hide"}
library(gdata)
library(caret)
library(randomForest)
```

## Load Datasets
```{r}
# Build training dataset
training <- read.csv("C://Users/vchieh/Downloads/pml-training.csv",header=TRUE, na.strings="NA")
training <-training[,c(8:11,37:49,60:68,84:86,102,113:124,140,151:160)]

# Split training set into 2 to cut down processing time
# Can also use model from 1st set to predict 2nd set
set.seed(123)
inTrain <- createDataPartition(y=training$classe,p=0.4,list=FALSE)
training1 <- training[inTrain,]
training2 <- training[-inTrain,]

# Build testing dataset
testing <- read.csv("C://Users/vchieh/Downloads/pml-testing.csv",header=TRUE, na.strings="NA")
testing <-testing[,c(8:11,37:49,60:68,84:86,102,113:124,140,151:160)]
```

Training and testing sets are subsetting in certain columns that have appropriate values to allow us to perform predictions.

## Analysis

```{r}
# Check for variance (remove if almost 0 since it does not affect the model too much, to reduce processing times)
variance <- data.frame(nrow=1,ncol=52)
for(i in 1:52){
  variance[1,i] <- var(training1[,i])
  names(variance)[i] <- names(training1)[i]
}
```

Most variances are not 0 or close to 0. We will keep all the measurements to save time in transforming the dataset.

## Build Model

```{r, eval=FALSE}
# Parallel Core Processing
library(doParallel)
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```

Parallel Core Processing done to speed up the process of training the model.

In this case, we use the random forests method to build a prediction model.

```{r}
# Define training control (Cross Validation with Random Forest method 'oob')
trControlRF <- trainControl(method="oob", number=3, repeats=1)

# Train the Model
modFit <- train(classe~ .,data=training1, ntree=500, method="rf", trControl = trControlRF)

# Make Predictions on 2nd training set
predicted <- predict(modFit, newdata=training2)

# Summarize Results on 2nd training set
confusionMatrix(predicted, training2$classe)
```

The model has a very high out-of-sample accuracy of 98.64% with an out-of-sample error of 1.36%, and we will use the model to make predictions on the test set.

```{r}
# Make Predictions on test set
answers <- predict(modFit, newdata=testing)
```

## End
```{r, eval=FALSE}
# Terminate Multiple Processes
stopCluster(cl)
registerDoSEQ()
```